% the acknowledgments section

\newthought{Team work makes the dream work.} I have worked towards this thesis over the past 9 years at Erasmus MC, and in that time have had the distinct pleasure of working with many amazing people. A big thank you to everybody who was part of my journey, thanks to all my co-authors from around the world, and I would like to thank the following people in particular:

\textbf{Dr Stubbs, dear Andrew.} Your guidance over the past 9 years shaped me into the scientist I am today. It is truly a pleasure working with you; thank you for your patience, easy-going demeanor, low-stress working environment, the many laughs, and for giving me the space to find my own path and building my confidence.

\textbf{Prof. Jenster, dear Guido.} Thank you for your guidance during my first years at EMC. You were never too busy to help when I had questions and always happy to explain the biology when I felt lost. Thanks for ending every email with a reminder to have fun, I needed it sometimes, but overall I succeeded and had a great deal of fun during my time at EMC.

%\textbf{Prof. van der Spek, dear Peter.} I would like to thank you for your guidance during the final phase of my promotion, your useful feedback and suggestions, as well as the help in navigating the practicalities around my promotion during the pandemic are greatly appreciated.

\textbf{Prof. van der Spek, other committee members.} Thanks for taking the time to carefully read my manuscript and providing insightful feedback that helped improve this thesis.

\textbf{Coworkers.} Thanks to everybody at the EMC for the years of fun, drinks, laughs, commiserations, rants, discussions, movie nights, dinners, even during a year of lockdown, it's been a blast!

\textbf{Galaxy community.} Thanks to everybody in the Galaxy community for making me feel welcome, and patiently helping me whenever I had questions. I learned lot from contributing to this project, and I've come to love the Galaxy community so much. A special thanks to Bérénice Batut, it's been a pleasure building the GTN and watching it grow over the years. Big shout out to Dave Clements for being the friendly face of an awesome community, you've been a great help and are always delightful to talk you.

\textbf{Women in Science} I feel lucky to work with so many amazing and strong women in science; Helena, Bérénice, Ines, Vera, Leslie, Subina, Beatriz, Wendi, Jen, Delphine, Anna, Anne, Maria, Elizabeth and many more, you are an inspiration and I've learned a bunch from all of you!

\textbf{Family.} Thanks to my Mom and Dad and two amazing sisters Miriam and Yvonne for always cheering me on, I love you all.

\textbf{Helena.} Last but not least, thanks to my amazing partner Helena for always being there for me, thanks for always reminding me that I can do this and having my back. Thanks for all the valuable discussions when I was writing this manuscript, and for co-authoring some of the papers in this thesis. Thanks also for helping me with the design of this thesis and your many opinions on fonts. Thanks for being my team mate, I love you.
