
\chapter{Curriculum Vitae}
\label{AppendixB}
\vspace{-1cm}
Saskia Hiltemann was born on July 24, 1984, in Tilburg, The Netherlands. She completed high school (VWO) in 2002 at the Alfrink College in Zoetermeer. After a life-long fascination with the stars, she then started a bachelor in Physics and Astronomy at the University of Leiden. During her bachelor she took some programming classes and loved it so much that she switched to a Bachelor in Computer Science, also at Leiden University, which she completed in 2008. While she was passionate about computer science, she missed the science and math she had come to love when studying Astronomy, so she chose to pursue a Master's degree in Bioinformatics. There she focused heavily on the theoretical side of bioinformatics, with her Master's thesis on the topic of cellular automata and reaction-diffusion systems. Saskia also always had a love for teaching; all throughout her studies she worked as a tutor of high school students, helping them with math and science. After completing her Master's degree, she started working at the Erasmus Medical Center, on various projects dedicated to building the infrastructure needed to analyse and interpret biological data, with a special focus on prostate cancer and later microbiome analysis. During her time at Erasmus MC she discovered a passion for the world of open-source software, and has contributed to many open source projects. She also co-founded the Galaxy Training Network, bringing together her love for teaching, science, and open source software.
