\chapter{List of publications}
\label{AppendixD}

Publications \ref{circos}, \ref{gmt}, \ref{gtn}, \ref{galaxy}, \ref{MYcrobiota}, \ref{VN}, \ref{ireport}, \ref{cgtag}, \ref{chromothripsis}, \ref{ifuse} are part of this thesis.

\begin{enumerate}
\setcounter{enumi}{-1}

\item Astrid Heikema, Rick Jansen, Saskia Hiltemann, John Hays, Andrew Stubbs, {\color{chaptergrey}{WeFaceNano: a user-friendly pipeline for complete ONT sequence assembly and detection of antibiotic resistance in multi-plasmid bacterial isolates}}, \textit{BMC Microbiology}, [in revision]

\item  Beatriz Serrano-Solano, Anika Erxleben, Cristóbal Gallardo-Alba, Helena Rasche, \textbf{Saskia Hiltemann}, Melanie Föll, Matthias Fahrner, Mark J. Dunning, Marcel Schulz, Beáta Scholtz, Dave Clements, Anton Nekrutenko, Bérénice Batut, Björn Grüning, {\color{chaptergrey}{Fostering Accessible Online Education Using Galaxy as an e-learning Platform}}, \textit{PLOS Computational Biology}, [accepted, March 2021]

\item Subina Mehta, Marie Crane, Emma Leith, Bérénice Batut, \textbf{Saskia Hiltemann}, Magnus Ø Arntzen, Benoit J. Kunath, Francesco Delogu, Ray Sajulga, Praveen Kumar, James E. Johnson, Timothy J. Griffin1, Pratik D. Jagtap {\color{chaptergrey}{ASaiM-MT: a validated and optimized ASaiM workflow for metatranscriptomics analysis within Galaxy framework}}, \textit{F1000}, February 2021, doi: 10.12688/f1000research.28608.1

\item  Willem de Koning, Milad Miladi, \textbf{Saskia Hiltemann}, Astrid Heikema, John P Hays, Stephan Flemming, Marius van den Beek, Dana A Mustafa, Rolf Backofen, Björn Grüning, Andrew P Stubbs {\color{chaptergrey}{NanoGalaxy: Nanopore long-read sequencing data analysis in Galaxy}}, \textit{GigaScience}, October 2020, doi: 10.1093/gigascience/giaa105 \label{one}

\item Astrid Heikema, Deborah Horst-Kreft, Stefan Boers, Rick Jansen, \textbf{Saskia Hiltemann}, Willem de Koning, Robert Kraaij, Maria de Ridder, Chantal van Houten, Louis Bont, Andrew  Stubbs and John Hays. {\color{chaptergrey}{Comparison of Illumina versus Nanopore 16S rRNAGene Sequencing of the Human Nasal Microbiota}}, \textit{Genes}, September 2020, doi: 10.3390/genes11091105

\item Helena Rasche, \textbf{Saskia Hiltemann}. {\color{chaptergrey}{Galactic Circos: User-friendly Circos plots within the Galaxy platform.}} \textit{GigaScience}, June 2020, doi: 10.1093/gigascience/giaa065 \label{circos}

\item \textbf{Saskia Hiltemann}, Stefan A Boers, Peter J van der Spek, Ruud Jansen, John P Hays, Andrew P Stubbs {\color{chaptergrey}{Galaxy mothur Toolset (GmT): a user-friendly application for 16S rRNA gene sequencing analysis using mothur.}} \textit{GigaScience}, February 2019, doi: 10.1093/gigascience/giy166 \label{gmt}


\item Bérénice Batut, \textbf{Saskia Hiltemann}, Andrea Bagnacani, Dannon Baker, Vivek Bhardwaj, Clemens Blank, Anthony Bretaudeau, Loraine Guéguen, Martin Čech, John Chilton, Dave Clements, Olivia Doppelt-Azeroual, Anika Erxleben, Mallory Freeberg, Simon Gladman, Youri Hoogstrate, Hans-Rudolf Hotz, Torsten Houwaart, Pratik Jagtap, Delphine Lariviere, Gildas Le Corguillé, Thomas Manke, Fabien Mareuil, Fidel Ramírez, Devon Ryan, Florian Sigloch, Nicola Soranzo, Joachim Wolff, Pavankumar Videm, Markus Wolfien, Aisanjiang Wubuli, Dilmurat Yusuf, Rolf Backofen, Anton Nekrutenko, Björn Grüning {\color{chaptergrey}{Community-driven data analysis training for biology.}}, June 2018, doi: 10.1016/j.cels.2018.05.012 \label{gtn}

\item Bérénice Batut, Kévin Gravouil, Clémence Defois, \textbf{Saskia Hiltemann}, Jean-François Brugère, Eric Peyretaillade, Pierre Peyret {\color{chaptergrey}{ASaiM: a Galaxy-based framework to analyze raw shotgun data from microbiota.}} \textit{GigaScience}, June 2018, doi: 10.1093/gigascience/giy057

\item  Enis Afgan, Dannon Baker, Bérénice Batut, Marius van den Beek, Dave Bouvier, Martin Čech, John Chilton, Dave Clements, Nate Coraor, Björn A Grüning, Aysam Guerler, Jennifer Hillman-Jackson, \textbf{Saskia Hiltemann}, Vahid Jalili, Helena Rasche, Nicola Soranzo, Jeremy Goecks, James Taylor, Anton Nekrutenko, Daniel Blankenberg {\color{chaptergrey}{The Galaxy platform for accessible, reproducible and collaborative biomedical analyses: 2018 update }}, \textit{Nucleic Acids Research}, May 2018, doi: 10.1093/nar/gky379 \label{galaxy}

\item Stefan A. Boers, \textbf{Saskia D. Hiltemann}, Andrew P. Stubbs, Ruud Jansen and John P. Hays. {\color{chaptergrey}{Development and evaluation of a culture-free microbiota profiling platform (MYcrobiota) for clinical diagnostics}}, March 2018, doi: 10.1007/s10096-018-3220-z \label{MYcrobiota}

\item Björn Grüning, Ryan  Dale, Andreas  Sjödin, Brad  Chapman, Jillian  Rowe, Christopher  Tomkins-Tinch, Renan  Valieris, Adam  Caprez, Bérénice  Batut, Mathias  Haudgaard, Thomas  Cokelaer, Kyle  Beauchamp, Brent  Pedersen, Youri  Hoogstrate, Anthony  Bretaudeau, Devon  Ryan, Gildas  Le Corguillé, Dilmurat  Yusuf, Sebastian  Luna-Valero, Rory  Kirchner, Karel  Brinda, Thomas  Wollmann, Martin  Raden, Simon  van Heeringen, Nicola  Soranzo, Lorena  Pantano, Zachary  Charlop-Powers, Per  Unneberg, Matthias  De Smet, Marcel  Martin, Greg  Von Kuster, Tiago  Antao, Milad  Miladi, Kevin  Thornton, Christian LU   Brueffer, Marius  van der Beek, Daniel  Maticzka, Clemens  Blank, Sebastian  Will, Kévin  Gravouil, Joachim  Wolff, Manuel  Holtgrewe, Jörg  Fallmann, Vitor  Piro, Ilya  Shlyakhter, Ayman  Yousif, Philip  Mabon, Xiao-Ou  Zhang, Wei  Shen, Jennifer  Cabral, Cristel  Thomas, Eric  Enns, Joseph  Brown, Jorrit  Boekel, Mattias  de Hollander, Jerome  Kelleher, Nitesh  Turaga, Julian  de Ruiter, Dave  Bouvier, Simon  Gladman, Saket  Choudhary, Nicholas  Harding, Florian  Eggenhofer, Arne  Kratz, Zhuoqing  Fang, Robert  Kleinkauf, Henning  Timm, Peter  Cock, Enrico  Seiler, Colin  Brislawn, Hai  Nguyen, Endre  Bakken Stovner, Philip  Ewels, Matt  Chambers, James  Johnson, Emil  Hägglund, Simon  Ye, Roman  Valls Guimera, Elmar  Pruesse, Augustine  Dunn, Lance  Parsons, Rob  Patro, David  Koppstein, Elena  Grassi, Inken  Wohlers, Alex  Reynolds, MacIntosh  Cornwell, Nicholas  Stoler, Daniel  Blankenberg, Guowei  He, Marcel  Bargull, Alexander  Junge, Rick  Farouni, Mallory  Freeberg, Sourav  Singh, Daniel  Bogema, Fabio  Cumbo, Liang-Bo  Wang, David  Larson, Matthew  Workentine, Upendra  Kumar Devisetty, Sacha  Laurent, Pierrick  Roger, Xavier  Garnier, Rasmus  Agren, Aziz  Khan, John  Eppley, Wei  Li, Bianca Katharina  Stöcker, Tobias  Rausch, James  Taylor, Patrick  Wright, Adam  Taranto, Davide  Chicco, Bengt  Sennblad, Jasmijn  Baaijens, Matthew  Gopez, Nezar  Abdennur, Iain  Milne, Jens  Preussner, Luca  Pinello, Avi  Srivastava, Aroon  Chande, Philip Reiner  Kensche, Yuri  Pirola, Michael  Knudsen, Ino  de Bruijn, Kai  Blin, Giorgio  Gonnella, Oana  Enache, Vivek  Rai, Nicholas  Waters, \textbf{Saskia Hiltemann}, Matthew Bendall, Christoph  Stahl, Alistair  Miles, Yannick  Boursin, Yasset  Perez-Riverol, Sebastian  Schmeier, Erik  Clarke, Kevin  Arvai, Matthieu  Jung, Tomás  Di Domenico, Julien  Seiler, Helena  Rasche, Etienne  Kornobis, Daniela  Beisser, Sven  Rahmann, Alexander  Mikheyev, Camy  Tran, Jordi  Capellades, Christopher  Schröder, Adrian Emanuel  Salatino, Simon  Dirmeier, Timothy  Webster, Oleksandr  Moskalenko, Stephen, Gordon and Köster, Johannes {\color{chaptergrey}{Bioconda: A sustainable and comprehensive software distribution for the life sciences.}}, \textit{Nature Methods}, July 2018, doi: 10.1038/s41592-018-0046-7

\item Chao Zhang, Jochem Bijlard, Christine Staiger, Serena Scollen4, David van Enckevort, Youri Hoogstrate, Alexander Senf, \textbf{Saskia Hiltemann}, Susanna Repo, Wibo Pipping, Mariska Bierkens, Stefan Payralbe, Bas Stringer, Jaap Heringa, Andrew Stubbs, Luiz Olavo Bonino Da Silva Santos, Jeroen Belien, Ward Weistra, Rita Azevedo, Kees van Bochove, Gerrit Meijer, Jan-Willem Boiten, Jordi Rambla, Remond Fijneman, J. Dylan Spalding, Sanne Abeln. {\color{chaptergrey}{Systematically linking tranSMART, Galaxy and EGA for reusing human translational research data.}}  \textit{F1000Research}, August 2017, doi: 10.12688/f1000research.12168.1.

\item Youri Hoogstrate, Chao Zhang, Alexander Senf, Jochem Bijlard, \textbf{Saskia Hiltemann}, David van Enckevort, Susanna Repo, Jaap Heringa, Guido Jenster, Remond J.A. Fijneman, Jan-Willem Boiten, Gerrit A. Meijer, Andrew Stubbs, Jordi Rambla, Dylan Spalding and Sanne Abeln. {\color{chaptergrey}{Integration of EGA secure data access into Galaxy.}}  \textit{F1000Research.} December 2016, doi: 10.12688/f1000research.10221.1.

\item Youri Hoogstrate, Rene Böttcher, \textbf{Saskia Hiltemann}, Peter van der Spek, Guido Jenster, Andrew Stubbs. {\color{chaptergrey}{FuMa: reporting overlap in RNA-seq detected fusion genes.}} \textit{Bioinformatics}, April 2016, doi: 10.1093/bioinformatics/btv721.

\item \textbf{Saskia Hiltemann}, Guido Jenster, Jan Trapman, Peter van der Spek, Andrew Stubbs. {\color{chaptergrey}{Discriminating somatic and germline mutations in tumor DNA samples without matching normals.}} \textit{Genome Research}, September 2015, doi: 10.1101/gr.183053.114. \label{VN}

\item Michael Moorhouse, David van Zessen, Hanna IJspeert, \textbf{Saskia Hiltemann}, Bas Horsman, Peter van der Spek, Mirjam van der Burg, Andrew Stubbs. {\color{chaptergrey}{ImmunoGlobulin galaxy (IGGalaxy) for simple determination and quantitation of immunoglobulin heavy chain rearrangements from NGS.}} \textit{BMC Immunology}, December 2014, doi: 10.1186/s12865-014-0059-7.

\item \textbf{Saskia Hiltemann}, Youri Hoogstrate, Peter van der Spek, Guido Jenster, Andrew Stubbs. {\color{chaptergrey}{iReport: a generalised Galaxy solution for integrated experimental reporting.}} \textit{GigaScience}, October 2014, doi: 10.1186/2047-217X-3-19. \label{ireport}

\item Ines Teles Alves, Thomas Hartjes, Elizabeth McClellan, \textbf{Saskia Hiltemann}, Rene Böttcher, Dits N, Temanni MR, Janssen B, van Workum W, Peter van der Spek, Andrew Stubbs, de Klein A, Eussen B, Jan Trapman, Guido Jenster. {\color{chaptergrey}{Next-generation sequencing reveals novel rare fusion events with functional implication in prostate cancer.}} \textit{Oncogene}, January 2015, doi: 10.1038/onc.2013.591.

\item \textbf{Saskia Hiltemann}, Hailiang Mei, Matthias de Hollander, Ivo Palli, Peter van der Spek, Guido Jenster, Andrew Stubbs. {\color{chaptergrey}{CGtag: complete genomics toolkit and annotation in a cloud-based Galaxy.}} \textit{GigaScience}, January 2014, doi: 10.1186/2047-217X-3-1. \label{cgtag}

\item \textbf{Saskia Hiltemann}, Elizabeth McClellan, Jos van Nijnatten, Bas Horsman, Ivo Palli, Ines Teles Alves, Thomas Hartjes, Jan Trapman, Peter van der Spek, Guido Jenster, Andrew Stubbs. {\color{chaptergrey}{iFUSE: integrated fusion gene explorer.}} \textit{Bioinformatics}, July 2013, doi: 10.1093/bioinformatics/btt252. \label{ifuse}

\item Ines Teles Alves, \textbf{Saskia Hiltemann}, Thomas Hartjes, Peter van der Spek, Andrew Stubbs, Jan Trapman, Guido Jenster. {\color{chaptergrey}{Gene fusions by chromothripsis of chromosome 5q in the VCaP prostate cancer cell line.}} \textit{Human Genetics}, June 2013, doi: 10.1007/s00439-013-1308-1. \label{chromothripsis}

\item Andrew Stubbs, Elizabeth McClellan, Bas Horsman, \textbf{Saskia Hiltemann }, Ivo Palli, Stephan Nouwens, Anton Koning, Hoogland F, Reumers J, Daphne Heijsman, Sigrid Swagemakers, Kremer A, Meijerink J, Lambrechts D, Peter van der Spek. {\color{chaptergrey}{Huvariome: a web server resource of whole genome next-generation sequencing allelic frequencies to aid in pathological candidate gene selection.}} \textit{Journal of Clinical Bioinformatics}, November 2012, doi: 10.1186/2043-9113-2-19.


\end{enumerate}
