%reset references counter
\setcounter{NAT@ctr}{-1}

\phantomsection
\addcontentsline{toc}{section}{Galaxy: the 2018 update}


\articletitle{Community-driven data analysis training \\ for biology}
Enis Afgan\textsuperscript{\ref{affil:hopkins},*},
Dannon Baker\textsuperscript{\ref{affil:hopkins}},
Bérénice Batut\textsuperscript{\ref{affil:freiburg},*},
Marius van den Beek\textsuperscript{\ref{affil:curie},*},
Dave Bouvier\textsuperscript{\ref{affil:pennstate},*},
Martin Čech\textsuperscript{\ref{affil:pennstate},*},
John Chilton\textsuperscript{\ref{affil:pennstate},*},
Dave Clements\textsuperscript{\ref{affil:hopkins},*},
Nate Coraor\textsuperscript{\ref{affil:pennstate},*},
Björn Grüning\textsuperscript{\ref{affil:freiburg,freiburg2},*},
Aysam Guerler\textsuperscript{\ref{affil:hopkins},*},
Jennifer Hillman-Jackson\textsuperscript{\ref{affil:pennstate},*},
Saskia Hiltemann\textsuperscript{\ref{affil:emc},*},
Vahid Jalili\textsuperscript{\ref{affil:oregon},*},
Eric Rasche\textsuperscript{\ref{affil:freiburg},*},
Nicola Soranzo\textsuperscript{\ref{affil:earlham}},
Jeremy Goecks\textsuperscript{\ref{affil:oregon},*},
James Taylor\textsuperscript{\ref{affil:hopkins},*},
Anton Nekrutenko\textsuperscript{\ref{affil:pennstate},*},
Daniel Blankenberg\textsuperscript{\ref{affil:cleveland},*}

* all authors contributed equally to this work.

\small
\begin{enumerate}
\itemsep-0.5em
\item Department of Biology, Johns Hopkins University, Baltimore MD USA. \label{affil:hopkins}
\item Department of Comuter Science, Albert-Ludwigs-University, Freiburg, Germany. \label{affil:freiburg}
\item Institut Curie, PSL Research University, Paris, France. \label{affil:curie}
\item Department of Biochemistry and Molecular Biology, Penn State University, University Park, PA, USA. \label{affil:pennstate}
\item Center for Biological Systems Analysis (ZBSA), University of Freiburg, Freiburg, Germany. \label{affil:freiburg2}
\item Department of Pathology, Erasmus Medical Centre, Rotterdam, The Netherlands. \label{affil:emc}
\item Department of Biomedical Engineering, Oregon Health and Science University, OR, USA.\label{affil:oregon}
\item Earlham Institute, Norwich Research Park, Norwich, United Kingdom. \label{affil:earlham}
\item Genomic Medicine Institute, Lerner Research Institute, Cleveland Clinic, Cleveland, OH, USA. \label{affil:cleveland}
\end{enumerate}
\normalsize

Published in: Nucleic Acid Research, April 2018 \\
DOI: \url{TODO } \\
%TODO: update doi when published

\section*{Abstract}

